# Simulating and fitting a 4 treatment experiment with correlated response variables

Let's simulate some data from a 4 treatment experiment. We will have 100 observations. 

```{r}
set.seed(42)
N <- 100
X <- matrix(c(
  c(rep(1, 25), rep(0, 75)),
  c(rep(0, 25), rep(1, 25), rep(0, 50)),
  c(rep(0, 50), rep(1, 25), rep(0, 25)),
  c(rep(0, 75), rep(1, 25))),
  ncol = 4)
head(X)
tail(X)
```

Let's create a correlation matrix and then calculate the covariance matrix given specified standard deviations. We will have 3 response variables.

```{r}
Omega <- rbind( # correlation matrix
  c(1, 0.5, 0.9),
  c(0.5, 1, 0.7),
  c(0.9, 0.7, 1)
)
sigma <- c(0.2, 0.3, 0.4) # residual SDs
Sigma <- diag(sigma) %*% Omega %*% diag(sigma) # covariance matrix
Sigma
Errors <- mvtnorm::rmvnorm(N, c(0,0,0), Sigma)
pairs(Errors)
round(cor(Errors), 2) # realized correlation
```

Now let's randomly generate some treatment effects or coefficients: 

```{r}
Betas <- round(matrix(runif(12, -1, 1), nrow = 3), 1)
Betas
```

And multiply our coefficients by the design matrix and add on the correlated errors to get our response measurements. 

```{r}
Y <- X %*% t(Betas) + Errors
```

Let's look at the data in various ways:

```{r, warning=FALSE, message=FALSE}
pairs(Y)
library(ggplot2)
library(dplyr)
d <- data.frame(x = c(rep(1, 25), rep(2, 25), rep(3, 25), rep(4, 25)), Y)
d <- reshape2::melt(d, id.vars = "x") %>% 
  mutate(variable = gsub("X", "Y", variable))
ggplot(d, aes(x, value)) + 
  geom_point(position = position_jitter(width = 0.1)) + 
  facet_wrap(~variable) +
  xlab("Treatment number")
```

Now let's fit our model with Stan:

```{r, warning=FALSE, message=FALSE}
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
m <- stan("multi.stan", 
  data = list(J = 4, K = 3, N = nrow(Y), x = X, y = Y),
  iter = 300)
print(m, pars = c("beta", "L_sigma", "Omega"), probs = c(0.1, 0.5, 0.9))
```

`beta` is the matrix of coefficients 
`Omega` are the elements of the correlation matrix
`L_sigma` contains the residual standard deviations

```{r}
plot(m, pars = "beta")
Betas
plot(m, pars = "Omega")
Omega
plot(m, pars = "L_sigma")
sigma
```
